---
title: "CD33 CAR NK: D1"
author: "Ahmad Al Ajami (AG Imkeller | KGU)"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: false
    toc: true
    toc_float: true
    toc_depth: 5
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, include=TRUE, message=FALSE, cache = FALSE}
suppressPackageStartupMessages({
library(tidyverse)
library(scran)
library(scater)
library(patchwork)
library(scDblFinder)
library(cowplot)
library(pheatmap)
library(edgeR)
library(ggrepel)
library(bluster)
})
```

# Loading Data

```{r data, message=FALSE}
data.dir <- "Files_already_generated_with_Sevenbridges/D1/"

data1 <- read.csv(paste(data.dir, "_1_Combined_Rhapsody19_Target1_RSEC_MolsPerCell.csv", sep = ""), skip = 7)
data1$Cell_Index <- paste(data1$Cell_Index, "_S1", sep = "")

data2 <- read.csv(paste(data.dir, "Combined_Rhapsody19_Target2_RSEC_MolsPerCell.csv", sep = ""), skip = 7)
data2$Cell_Index <- paste(data2$Cell_Index, "_S2", sep = "")

data <- rbind(data1, data2) 
```

The metadata contains Sample_Tag (nucleotide labeled antibody used to distinguish between samples), Sample_Name (patient id), multiplet status assigned by the SevenBridges pipeline

```{r metadata, message=FALSE}
# batch 1
## table with cell information (sample barcode, donor, ...)
coldata1 <- read.csv(paste(data.dir, "_1_Rhapsody19_Target1_Sample_Tag_Calls.csv", sep = ""), skip = 7) %>%
  dplyr::mutate(invalid = (Sample_Name == "Multiplet" | Sample_Name == "Undetermined")) %>%
  dplyr::mutate(batch = "S1")
coldata1$Cell_Index <- paste(coldata1$Cell_Index, "_S1", sep = "")

# batch 2
## table with cell information (sample barcode, donor, ...)
coldata2 <- read.csv(paste(data.dir, "Rhapsody19_Target2_Sample_Tag_Calls.csv", sep = ""), skip = 7) %>%
  dplyr::mutate(invalid = (Sample_Name == "Multiplet" | Sample_Name == "Undetermined")) %>%
  dplyr::mutate(batch = "S2")
coldata2$Cell_Index <- paste(coldata2$Cell_Index, "_S2", sep = "")

# binding both coldata from the 2 batches
coldata <- rbind(coldata1, coldata2)
```

## SevenBridges Output

SevenBridges pre-processing pipeline report:

```{r sb_output, message=FALSE}
sb <- data.frame(Batch = c("S1", "S2"),
                 Number.of.Cells = c(dim(data1)[1], dim(data2)[1]),
                 Number.of.Features = c(dim(data1[, -1])[2], dim(data2[, -1])[2]),
                 Median.Number.of.Features = c(median(rowSums(data1 > 0)), median(rowSums(data2 > 0))),
                 stringsAsFactors = F)

DT::datatable(sb, 
              class = 'compact stripe hower',
              options = list(
                dom = "Bfrtip",
                scrollX = TRUE,
                paging = FALSE,
                searching = FALSE,
                info = FALSE,
                ordering = FALSE,
                columnDefs = list(list(className = 'dt-center', targets = 0:3))), rownames = FALSE)
```

## Sample IDs

Number of cells in each sample:

```{r sample_ids, message=FALSE}
table(coldata$Sample_Name)
```

We define a new column containing the group of samples:

- BeforeCoc = Before
- AfterCoc = After
- NA = Multiplets + Undetermined

```{r group_ids, message=FALSE, fig.width=10, fig.height=4}
coldata <- coldata %>%
  dplyr::mutate(group = ifelse(grepl("before", Sample_Name), "BeforeCoc",
    ifelse(grepl("after", Sample_Name), "AfterCoc", NA)))

p1 <- coldata %>%
  ggplot(aes(x = batch, fill = group)) + 
  geom_bar() +
  labs(x = "Batch", fill = "Group")

p2 <- coldata %>%
  ggplot(aes(x = batch, fill = Sample_Name)) + 
  geom_bar() +
  labs(x = "Batch", fill = "Sample Name")

p3 <- coldata %>%
  ggplot(aes(x = group, fill = Sample_Name)) + 
  geom_bar() +
  labs(x = "Group", fill = "Sample Name")

p1 + p2 + p3
```

```{r}
coldata %>%
  ggplot(aes(x = batch, fill = group, color = Sample_Name)) +
  geom_bar() +
  labs(x = "Batch", fill = "Group", color = "Sample Name")
```

# Pre-processing Workflow

## Filtering

```{r, message=FALSE}
data_t <- data[, -1] %>% t()
rownames(data_t) <- colnames(data[, -1])
colnames(data_t) <- data$Cell_Index

sce <- SingleCellExperiment(assays = list(counts = data_t), colData = coldata)
rowData(sce)$Type <- ifelse(grepl("abo", rownames(sce), ignore.case = TRUE), "Antibody Capture", "Gene Expression") # 'Type' is either gene or antibody
```

```{r, message=FALSE}
# removing the cells with multiplet or undetermined assignment
sce <- sce[, !coldata$invalid]

coldata <- coldata %>%
  dplyr::filter(!invalid)
```

```{r, message=FALSE}
# splitting protein from RNA
sce_split <- splitAltExps(sce, rowData(sce)$Type)
# altExpNames(sce_split)
```

### Overall Quality Metrics

#### Before Filtering

```{r, message=FALSE}
# ADT vs RNA counts
rna_numi <- colSums(counts(sce_split))
adt_numi <- colSums(counts(altExp(sce_split)))
nUMIdf <- data.frame(rna_nUMI = rna_numi,
                     adt_nUMI = adt_numi, 
                     Batch = sce_split$batch)

ggplot(nUMIdf, aes(x = log10(adt_nUMI+1), y = log10(rna_nUMI+1))) + 
  geom_point(alpha = 0.1, size = 1) +
  geom_density_2d(color = "orange") +
  facet_grid(.~Batch) +
  labs(x = "ADT Counts", y = "RNA Counts")
```

#### Computing QC metrics

```{r scDblFinder, echo=TRUE, include=FALSE}
# detecting doublets using scDblFinder
# keeping the doublets returned and rechecking them in downstream analysis
set.seed(42L)

sce_split <- scDblFinder(sce_split)
```

```{r, message=FALSE}
outliers <- perCellQCMetrics(sce_split)
libsize.drop <- isOutlier(outliers$sum, log = TRUE, type = "both", nmads = 3)
feature.drop <- isOutlier(outliers$detected, log = TRUE, type = "lower", nmads = 3)
print(data.frame(ByLibSize = sum(libsize.drop),
                 ByFeature = sum(feature.drop),
                 Discard = sum(libsize.drop | feature.drop)))
outliers.discard <- libsize.drop | feature.drop

#outliers.discard <- quickPerCellQC(outliers)
#colSums(as.matrix(outliers.discard))
```

#### Ab Counts

```{r, message=FALSE}
ab.discard <- isOutlier(outliers$`altexps_Antibody Capture_detected`,
                        log = TRUE,
                        type = "lower",
                        min_diff = 1)
summary(ab.discard)

hist(outliers$`altexps_Antibody Capture_detected`, 
     col = 'grey',
     main = "", xlab = "Number of detected ADTs")
```

#### QC Plots

```{r, message=FALSE}
colData(sce_split) <- cbind(colData(sce_split), outliers) # adding the outliers info (sum, detected) to the sce object coldata
sce_split$discard <- outliers.discard | ab.discard # adding the cells to be discarded to the sce object coldata

plotColData(sce_split, x = "detected", y = "sum", colour_by = "discard", other_fields=c("batch")) +
  facet_wrap(~batch) +
  theme(panel.border = element_rect(color = "grey")) + 
  labs(x = "Number of Features Detected", y = "Library Size")
```

```{r, message=FALSE}
gridExtra::grid.arrange(
  plotColData(sce_split, x="batch", y="sum", colour_by="discard") +
    scale_y_log10() + ggtitle("Library Size"),
  plotColData(sce_split, x="batch", y="detected", colour_by="discard") +
    scale_y_log10() + ggtitle("Genes Detected") + labs(y = "genes detected"),
  plotColData(sce_split, x="batch", y="altexps_Antibody Capture_detected", colour_by="discard") +
    scale_y_log10() + ggtitle("Antibodies Detected") + labs(y = "antibodies detected"),
  ncol=1
)
```

```{r, message=FALSE}
gridExtra::grid.arrange(
  plotColData(sce_split, y="sum", colour_by="discard") +
    scale_y_log10() + ggtitle("Library Size"),
  plotColData(sce_split, y="detected", colour_by="discard") +
    scale_y_log10() + ggtitle("Genes Detected") + labs(y = "genes detected"),
  plotColData(sce_split, y="altexps_Antibody Capture_detected", colour_by="discard") +
    scale_y_log10() + ggtitle("Antibodies Detected") + labs(y = "antibodies detected"),
  ncol=1
)
```

```{r, message=FALSE}
# ADT vs RNA counts
rna_numi <- colSums(counts(sce_split))
adt_numi <- colSums(counts(altExp(sce_split)))
nUMIdf <- data.frame(rna_nUMI = rna_numi,
                     adt_nUMI = adt_numi, 
                     Batch = sce_split$batch,
                     Discard = sce_split$discard)

ggplot(nUMIdf, aes(x = log10(adt_nUMI+1), y = log10(rna_nUMI+1), color = Discard)) + 
  geom_point(alpha = 0.1, size = 1) +
  geom_density_2d(color = "black") +
  facet_grid(.~Batch) +
  labs(x = "ADT Counts", y = "RNA Counts") +
  scale_color_manual(values = c("#0072B2", "#C4961A"))
```

```{r filtering, message=FALSE}
# filtering; removing outliers
sce_split <- sce_split[, !sce_split$discard]
sce_split
table(coldata$Sample_Name)
```

## Normalization

```{r, message=FALSE}
# RNA

## size factors
lib.sf <- librarySizeFactors(sce_split)
summary(lib.sf)

lib.sf.df <- data.frame(lib.sf, Group = colData(sce_split)$group)

## plotting size factors
hist(log10(lib.sf), xlab="Log10[Size Factor]", col='grey80')

ggplot(lib.sf.df, aes(log10(lib.sf))) +  
  geom_histogram() + 
  facet_wrap(~Group, scales = "free") + 
  labs(x = "Log10[Size Factor]")

## normalization
sce_split <- logNormCounts(sce_split)
#assayNames(sce_split)

# ADT

clr_norm <- function(x) {
  return(log1p(x = x / (exp(x = sum(log1p(x = x[x > 0]), na.rm = TRUE) / length(x = x)))))
}

logcounts(altExp(sce_split)) <- apply(counts(altExp(sce_split)), 2, clr_norm)
```

## Variance Modeling

```{r, message=FALSE}
# determining variance components
dec <- modelGeneVar(sce_split)
dec[order(dec$bio, decreasing = TRUE), ]

fit <- metadata(dec)
plot(fit$mean, fit$var, 
     xlab = "Mean of log-expression", ylab = "Variance of log-expression")
points(fit$mean, fit$var, col = "red", pch = 16)
curve(fit$trend(x), col = "dodgerblue", add = TRUE, lwd = 2)
```

## Distribution of ADT Markers 

Looking at the distribution of ADT markers (after normalization): 

```{r, message=FALSE, fig.asp=1.8}
p <- do.call(plot_grid, c(lapply(rownames(logcounts(altExp(sce_split))), function(adt_name){
  ggplot(as.data.frame(logcounts(altExp(sce_split))[adt_name, ]), aes_string(x=logcounts(altExp(sce_split))[adt_name, ])) +
    geom_histogram(color="black",
                   fill = "black",
                   breaks=seq(0, 4.5, by=0.10)) + 
    xlab(adt_name) + 
    theme(axis.title.x = element_text(size = 8, face="bold"))}), ncol = 4))

print(p)
```

We define a new column containing the condition of samples:

- NT = NTbefore + NTafter
- KO = KObefore + KOafter
- CAR = CARbefore + CARafter
- CARKO = CARKObefore + CARKOafter

```{r, message=FALSE}
# creating new colData
sce_split$CAR <- ifelse(sce_split$Sample_Name %in% c("CARKOafter", "CARKObefore", "CARafter", "CARbefore"), "Yes", "No")
sce_split$KO <- ifelse(sce_split$Sample_Name %in% c("CARKOafter", "CARKObefore", "KOafter", "KObefore"), "Yes", "No")
sce_split$condition <- ifelse(sce_split$Sample_Name %in% c("CARKOafter", "CARKObefore"), "CARKO",
                                 ifelse(sce_split$Sample_Name %in% c("CARafter", "CARbefore"), "CAR",
                                        ifelse(sce_split$Sample_Name %in% c("KOafter", "KObefore"), "KO", "NT")))
sce_split$condition <- factor(sce_split$condition, levels = c("NT", "KO", "CAR", "CARKO"))
```

```{r}
saveRDS(sce_split, "intermediate_files/D1/sce_donor1.Rds")
```

# Processing - Downstream Analysis

## Before

```{r}
sce_split <- readRDS("intermediate_files/D1/sce_donor1.Rds")
sce_before <- sce_split[, colData(sce_split)$group == "BeforeCoc"]
sce_counts <- rbind(counts(sce_before), counts(altExp(sce_before)))
sce_logcounts <- rbind(logcounts(sce_before), logcounts(altExp(sce_before)))
sce <- SingleCellExperiment(assays = list(counts = sce_counts, logcounts = sce_logcounts))
rowData(sce)$Type <- ifelse(grepl("abo", rownames(sce), ignore.case = TRUE), "Antibody Capture", "Gene Expression")
sce$condition <- sce_before$condition
sce$condition <- gsub("NT", "NT-NK", sce$condition)
sce$condition <- gsub("CAR", "CAR33-NK", sce$condition)
sce$condition <- gsub("KO", "KLRC1.ko-NK", sce$condition)
sce$condition <- gsub("CAR33-NKKLRC1.ko-NK", "CAR33-KLRC1.ko-NK", sce$condition)
sce$condition <- factor(sce$condition, levels = c("NT-NK", "KLRC1.ko-NK", "CAR33-NK", "CAR33-KLRC1.ko-NK"))
```

Start - supp figure 2

```{r}
sce_v2 <- sce
rownames(sce_v2) <- gsub("TCR.alpha_beta.TRA_TRB.AHS0078.pAbO", "TCR_alpha_beta.TRA_TRB.AHS0078.pAbO", rownames(sce_v2))
rownames(sce_v2[grep("abo", rownames(sce_v2), ignore.case = T, value = T), ]) <- 
  gsub(".AH.*", "", rownames(sce_v2[grep("abo", rownames(sce_v2), ignore.case = T, value = T), ]))

sce_supp <- splitAltExps(sce_v2, rowData(sce_v2)$Type)
rownames(altExp(sce_supp)) <- gsub("\\..*", "", rownames(altExp(sce_supp)))


adts <- c("CD3", "CD4", "TCR_alpha_beta")

supp2 <- do.call(plot_grid, c(lapply(adts, function(adt_name){
  ggplot(as.data.frame(logcounts(altExp(sce_supp))[adt_name, ]), 
         aes_string(x=logcounts(altExp(sce_supp))[adt_name, ])) +
    geom_histogram(color="black",
                   fill = "black",
                   breaks=seq(0, 4.5, by=0.10)) + 
    xlab(adt_name) + 
    theme(axis.title.x = element_text(size = 8, face="bold"))}), ncol = 3))

ggsave("fig2supp1.tiff", supp2, width = 6, height = 2)
ggsave("fig2supp1.pdf", supp2, width = 6, height = 2)
```

End - supp figure 2

```{r}
'%nin%' <- Negate('%in%')
sce <- sce[rownames(sce) %nin% c("CD3.SK7.CD3E.AHS0033.pAbO", "CD4.SK3.CD4.AHS0032.pAbO", "TCR.alpha_beta.TRA_TRB.AHS0078.pAbO", 
                                 "TCR.gamma_delta.11F2.TRG_TRD.AHS0142.pAbO", "IgG.IGHG1_2_3_4.AHS0059.pAbO", "IgM.IGHM.AHS0198.pAbO"), ]
```


```{r}
set.seed(42L)

sce <- runPCA(sce, ncomponents=20)
sce <- runTSNE(sce, dimred="PCA")

snn_graph <- buildSNNGraph(sce, k=50, use.dimred='PCA')
igraph_clusters <- igraph::cluster_louvain(snn_graph)$membership
sce$lbls_50 <- as.factor(igraph_clusters)

snn_graph <- buildSNNGraph(sce, k=70, use.dimred='PCA')
igraph_clusters <- igraph::cluster_louvain(snn_graph)$membership
sce$lbls_70 <- as.factor(igraph_clusters)

snn_graph <- buildSNNGraph(sce, k=100, use.dimred='PCA')
igraph_clusters <- igraph::cluster_louvain(snn_graph)$membership
sce$lbls_100 <- as.factor(igraph_clusters)

table(lbls_50 = sce$lbls_50, lbls_70 = sce$lbls_70)
table(lbls_50 = sce$lbls_50, lbls_100 = sce$lbls_100)
table(lbls_70 = sce$lbls_70, lbls_100 = sce$lbls_100)
```

```{r}
k_param <- c("lbls_50","lbls_70","lbls_100")
m1 <- lapply(k_param, function(.){
  print(.)
  sil.approx <- approxSilhouette(reducedDim(sce, "PCA"), clusters=colData(sce)[[.]])
  sil.data <- as.data.frame(sil.approx)
  sil.data$closest <- factor(ifelse(sil.data$width > 0, colData(sce)[[.]], sil.data$other))
  sil.data$cluster <- colData(sce)[[.]]
  ari <- aricode::ARI(sil.data$cluster, sil.data$closest)
  wri <- Wind::wRI(sil.data$cluster, sil.data$closest)
  plot(ggplot(sil.data, aes(x=cluster, y=width, colour=closest)) +
    ggbeeswarm::geom_quasirandom(method="smiley"))
  return(list(ari, wri[[1]]))
})
df1 <- data.frame(k = c(50,70,100),
                  ARI = sapply(m1, function(x) x[[1]]),
                  WRI = sapply(m1, function(x) x[[2]]),
                  Method = "Silhouette Width")

m2 <- lapply(k_param, function(.){
  print(.)
  pure.pbmc <- neighborPurity(reducedDim(sce, "PCA"), colData(sce)[[.]])
  pure.data <- as.data.frame(pure.pbmc)
  pure.data$maximum <- factor(pure.data$maximum)
  pure.data$cluster <- colData(sce)[[.]]
  ari <- aricode::ARI(pure.data$cluster, pure.data$maximum)
  wri <- Wind::wRI(pure.data$cluster, pure.data$maximum)
  plot(ggplot(pure.data, aes(x=cluster, y=purity, colour=maximum)) +
    ggbeeswarm::geom_quasirandom(method="smiley"))
  return(list(ari, wri[[1]]))
})
df2 <- data.frame(k = c(50,70,100),
                  ARI = sapply(m2, function(x) x[[1]]),
                  WRI = sapply(m2, function(x) x[[2]]),
                  Method = "Cluster Purity")

df <- rbind(df1, df2)
df <- pivot_longer(df, cols = c("ARI", "WRI"), names_to = "score", values_to = "Value")
p <- ggplot(df, aes(x = as.factor(k), y = Value, fill = score)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~Method, ncol = 1) +
  labs(x = "k", y = "similarity score") +
  scale_fill_manual(values = c("#1B9E77", "#D95F02"))
p
ggsave("fig3supp_clus.pdf", p, width = 4, height = 4)
ggsave("fig3supp_clus.tiff", p, width = 4, height = 4)
```

```{r}
sce$igraph_lbls <- sce$lbls_50
```

```{r}
p1 <- plotReducedDim(sce, dimred = "TSNE", by_exprs_values = "logcounts", colour_by = "condition", other_fields = "condition") +
  facet_wrap(~condition, ncol = 4) + 
  scale_color_manual(values = c("#a0a0a3", "#90bff9", "#f2b77d", "#ff8080")) +
  theme_bw() + 
  theme(strip.text.x = element_text(size = 12), legend.title = element_blank(), legend.position="bottom", legend.text = element_text(size=10))

p2 <- plotReducedDim(sce, dimred = "TSNE", by_exprs_values = "logcounts", colour_by = "igraph_lbls", other_fields = "igraph_lbls") +
  facet_wrap(~igraph_lbls, ncol = 6) + 
  scale_color_manual(values = c("green", "orange", "purple", "navy", "cyan", "maroon")) +
  theme_bw() + 
  theme(strip.text.x = element_text(size = 12), legend.position="bottom", legend.text = element_text(size=10)) +
  guides(colour = guide_legend(nrow = 1)) +
  labs(color = "cluster")

plot_grid(p1 , p2, ncol = 1, align = "hv", labels = c("A", "B"))
ggsave("fig3a.tiff", p1, height = 2.5, width = 7.5)
ggsave("fig3c.tiff", p2, width = 7, height = 2)

ggsave("fig3a.pdf", p1, height = 2.5, width = 7.5)
ggsave("fig3c.pdf", p2, width = 7, height = 2)
```

```{r, fig.width=6, fig.height=9}
p3 <- data.frame(colData(sce)) %>%
  group_by(igraph_lbls) %>%
  summarize(n = n()) %>%
  mutate(freq = n / sum(n) * 100) %>% 
  ggplot(aes(x = igraph_lbls, y = freq, fill = igraph_lbls)) +
  geom_col(position = "dodge", color = "black") + 
  scale_fill_manual(values = c("green", "orange", "purple", "navy", "cyan", "maroon"))+ 
  labs(x = "cluster", y = "cluster relative abundance (%)", fill = "cluster") +
  theme_bw() +
  theme(legend.position="bottom", legend.text = element_text(size=10)) +
  guides(fill = guide_legend(nrow = 1))

p4 <- data.frame(colData(sce)) %>%
  group_by(condition, igraph_lbls) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = igraph_lbls, y = n, fill = condition)) +
  geom_col(position = "fill", color = "black") + 
  scale_fill_manual(values = c("gray", "blue", "orange", "red")) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  labs(title = "Cluster Abundance", x = "Cluster Number", y = "Cluster Relative Abundance")

p5 <- data.frame(colData(sce)) %>%
  group_by(condition, igraph_lbls) %>%
  summarize(n = n()) %>%
  mutate(freq = n / sum(n)) %>% 
  ggplot(aes(x = igraph_lbls, y = freq, fill = condition)) +
  geom_col(position = "dodge", color = "black") + 
  scale_fill_manual(values = c("#a0a0a3", "#90bff9", "#f2b77d", "#ff8080")) + 
  labs(x = "cluster", y = "condition relative abundance") +
  theme_bw() +
  theme(legend.position="bottom", legend.text = element_text(size=10)) +
  guides(fill = guide_legend(nrow = 1))

plot_grid(p3, p4, p5, ncol = 1, align = "hv", labels = c("A", "B", "C"))
ggsave("fig3d.tiff", p3, height = 4, width = 4)
ggsave("fig3e.tiff", p5, height = 4, width=6)

ggsave("fig3d.pdf", p3, height = 4, width = 4)
ggsave("fig3e.pdf", p5, height = 4, width=6)
```

```{r}
rownames(sce) <- gsub("TCR.alpha_beta.TRA_TRB.AHS0078.pAbO", "TCR_alpha_beta.TRA_TRB.AHS0078.pAbO", rownames(sce))
rownames(sce) <- gsub("TCR.gamma_delta.11F2.TRG_TRD.AHS0142.pAbO", "TCR_gamma_delta.11F2.TRG_TRD.AHS0142.pAbO", rownames(sce))
rownames(sce[grep("abo", rownames(sce), ignore.case = T, value = T), ]) <- gsub(".AH.*", "", rownames(sce[grep("abo", rownames(sce), ignore.case = T, value = T), ]))
```

```{r}
sce_split <- splitAltExps(sce, rowData(sce)$Type)
rownames(altExp(sce_split)) <- gsub("\\..*", "", rownames(altExp(sce_split)))
altExp(sce_split)$igraph_lbls <- sce_split$igraph_lbls
```

```{r}
get_top_markers <- function(df) {
  top_genes <- as.data.frame(df) %>%
    dplyr::slice_min(order_by = Top, n = 8)
  return(top_genes)
}

markers_rna <- findMarkers(sce_split, groups = sce_split$igraph_lbls)
top_markers_rna <- lapply(markers_rna, get_top_markers)
top_markers_unique_rna <- unique(do.call(c, lapply(top_markers_rna, rownames)))

markers_adt <- findMarkers(altExp(sce_split), groups = altExp(sce_split)$igraph_lbls)
top_markers_adt <- lapply(markers_adt, get_top_markers)
top_markers_unique_adt <- unique(do.call(c, lapply(top_markers_adt, rownames)))

p1 <- plotDots(sce_split, features = top_markers_unique_rna, group = "igraph_lbls",
              color = c("darkblue", "lightblue", "gray", "yellow", "orange", "darkorange", "red", "darkred")) +
  theme(legend.position="bottom", legend.box="vertical", legend.margin=margin())

p2 <- plotDots(altExp(sce_split), features = top_markers_unique_adt, group = "igraph_lbls",
              color = c("darkblue", "lightblue", "gray", "yellow", "orange", "darkorange", "red", "darkred")) +
  theme(legend.position="bottom", legend.box="vertical", legend.margin=margin())

ggsave("fig3f1.tiff", p1, width = 6, height = 12)
ggsave("fig3f1.pdf", p1, width = 6, height = 12)

ggsave("fig3f2.tiff", p2, width = 6, height = 12)
ggsave("fig3f2.pdf", p2, width = 6, height = 12)
```

```{r}
altExp(sce_split)$CD16_status <- ifelse(logcounts(altExp(sce_split)["CD16",]) > 1, "CD16+", "CD16-")
altExp(sce_split)$condition <- sce_split$condition
```

```{r}
adts <- rownames(altExp(sce_split))[which(!rownames(altExp(sce_split)) %in% "CD16")]

supp_v1 <- lapply(adts, function(adt_name){
  ggplot(data.frame(value = logcounts(altExp(sce_split))[adt_name, ],
                    CD16_status = altExp(sce_split)$CD16_status),
         aes(x = value, color = CD16_status, fill = CD16_status)) +
    geom_density(alpha = 0.5) +
    scale_fill_manual(values = c("CD16-" = "lightblue", "CD16+" = "lightcoral")) + 
    scale_color_manual(values = c("CD16-" = "blue", "CD16+" = "red")) + 
    xlab(adt_name) + 
    theme(axis.title.x = element_text(size = 8, face="bold"),
          legend.position = "right")})

final_plot <- wrap_plots(supp_v1, ncol = 4) +
  plot_layout(guides = "collect")

ggsave("fig2.v1.tiff", final_plot, width = 12, height = 18)
ggsave("fig2.v1.pdf", final_plot, width = 12, height = 18)

# ------------------------------------- #

adts <- c("CD56", "CD69", "CD279", "CD18", "CD38", "CD94", "CD122", "CD226", "CD314", "CD335", "Tim3")

supp_v2 <- lapply(adts, function(adt_name){
  ggplot(data.frame(value = logcounts(altExp(sce_split))[adt_name, ],
                    #CD16_status = altExp(sce_split)$CD16_status,
                    condition = altExp(sce_split)$condition),
         # aes(x = value, color = condition, linetype = CD16_status)) +
         aes(x = value, color = condition)) +
    geom_density(alpha = 0.5) +
    scale_color_manual(values = c("NT-NK" = "gray", "KLRC1.ko-NK" = "blue", "CAR33-NK" = "orange", "CAR33-KLRC1.ko-NK" = "red")) +
    # scale_linetype_manual(values = c("CD16-" = "dashed", "CD16+" = "solid")) + 
    xlab(adt_name) + 
    theme(axis.title.x = element_text(size = 8, face="bold"),
          legend.position = "right")})

final_plot <- wrap_plots(supp_v2, ncol = 3) +
  plot_layout(guides = "collect")

ggsave("fig2.v2.tiff", final_plot, width = 10, height = 10)
ggsave("fig2.v2.pdf", final_plot, width = 10, height = 10)
```

Additional:

```{r}
supp1 <- plotDots(altExp(sce_split), features = rownames(altExp(sce_split)), group = "CD16_status",
                  color = c("darkblue", "lightblue", "gray", "yellow", "orange", "darkorange", "red", "darkred")) +
  theme(legend.position="bottom", legend.box="vertical", legend.margin=margin())

ggsave("fig3supp1.tiff", supp1, width = 6, height = 12)
ggsave("fig3supp1.pdf", supp1, width = 6, height = 12)
```

```{r}
supp1.1 <- do.call(plot_grid, c(lapply(rownames(logcounts(altExp(sce_split))), function(adt_name){
  ggplot(as.data.frame(logcounts(altExp(sce_split))[adt_name, altExp(sce_split)$CD16_status == "CD16+"]), 
         aes_string(x=logcounts(altExp(sce_split))[adt_name, altExp(sce_split)$CD16_status == "CD16+"])) +
    geom_histogram(color="black",
                   fill = "black",
                   breaks=seq(0, 4.5, by=0.10)) + 
    xlab(adt_name) + 
    theme(axis.title.x = element_text(size = 8, face="bold"))}), ncol = 4))

ggsave("fig3supp1.1.tiff", supp1.1, width = 8, height = 12)
ggsave("fig3supp1.1.pdf", supp1.1, width = 8, height = 12)

supp1.2 <- do.call(plot_grid, c(lapply(rownames(logcounts(altExp(sce_split))), function(adt_name){
  ggplot(as.data.frame(logcounts(altExp(sce_split))[adt_name, altExp(sce_split)$CD16_status == "CD16-"]), 
         aes_string(x=logcounts(altExp(sce_split))[adt_name, altExp(sce_split)$CD16_status == "CD16-"])) +
    geom_histogram(color="black",
                   fill = "black",
                   breaks=seq(0, 4.5, by=0.10)) + 
    xlab(adt_name) + 
    theme(axis.title.x = element_text(size = 8, face="bold"))}), ncol = 4))

ggsave("fig3supp1.2.tiff", supp1.2, width = 8, height = 12)
ggsave("fig3supp1.2.pdf", supp1.2, width = 8, height = 12)
```

```{r}
supp2.1 <- do.call(plot_grid, c(lapply(rownames(logcounts(altExp(sce_split))), function(adt_name){
  ggplot(as.data.frame(logcounts(altExp(sce_split))[adt_name, ]), 
         aes_string(x=logcounts(altExp(sce_split))[adt_name, ])) +
    geom_histogram(color="black",
                   fill = "black",
                   breaks=seq(0, 4.5, by=0.10)) + 
    xlab(adt_name) + 
    theme(axis.title.x = element_text(size = 8, face="bold"))}), ncol = 4))

ggsave("fig3supp2.1.tiff", supp2.1, width = 8, height = 12)
ggsave("fig3supp2.1.pdf", supp2.1, width = 8, height = 12)
```

```{r}
sce_split <- readRDS("intermediate_files/D1/sce_donor1.Rds")
sce_split$condition <- gsub("NT", "NT-NK", sce_split$condition)
sce_split$condition <- gsub("CAR", "CAR33-NK", sce_split$condition)
sce_split$condition <- gsub("KO", "KLRC1.ko-NK", sce_split$condition)
sce_split$condition <- gsub("CAR33-NKKLRC1.ko-NK", "CAR33-KLRC1.ko-NK", sce_split$condition)
sce_split$condition <- factor(sce_split$condition, levels = c("NT-NK", "KLRC1.ko-NK", "CAR33-NK", "CAR33-KLRC1.ko-NK"))

altExp(sce_split)$group <- sce_split$group
altExp(sce_split)$condition <- sce_split$condition

rownames(altExp(sce_split)) <- gsub("TCR.alpha_beta.TRA_TRB.AHS0078.pAbO", "TCR_alpha_beta.TRA_TRB.AHS0078.pAbO", rownames(altExp(sce_split)))
rownames(altExp(sce_split)) <- gsub("TCR.gamma_delta.11F2.TRG_TRD.AHS0142.pAbO", "TCR_gamma_delta.11F2.TRG_TRD.AHS0142.pAbO", rownames(altExp(sce_split)))
rownames(altExp(sce_split)) <- gsub("\\..*", "", rownames(altExp(sce_split)))

altExp(sce_split)$condition_group <- paste(altExp(sce_split)$condition, altExp(sce_split)$group, sep = "_")

supp3 <- plotDots(altExp(sce_split), features = rownames(altExp(sce_split)), group = "condition_group",
                  color = c("darkblue", "lightblue", "gray", "yellow", "orange", "darkorange", "red", "darkred")) +
  theme(legend.position="bottom", legend.box="vertical", legend.margin=margin())

ggsave("fig3supp3.tiff", supp3, width = 18, height = 12)
ggsave("fig3supp3.pdf", supp3, width = 18, height = 12)
```

```{r}
supp3.1 <- do.call(plot_grid, c(lapply(rownames(logcounts(altExp(sce_split))), function(adt_name){
  ggplot(as.data.frame(logcounts(altExp(sce_split))[adt_name, altExp(sce_split)$condition_group == "CAR33-KLRC1.ko-NK_AfterCoc"]), 
         aes_string(x=logcounts(altExp(sce_split))[adt_name, altExp(sce_split)$condition_group == "CAR33-KLRC1.ko-NK_AfterCoc"])) +
    geom_histogram(color="black",
                   fill = "black",
                   breaks=seq(0, 4.5, by=0.10)) + 
    xlab(adt_name) + 
    theme(axis.title.x = element_text(size = 8, face="bold"))}), ncol = 4))

ggsave("fig3supp3.1.tiff", supp3.1, width = 8, height = 12)
ggsave("fig3supp3.1.pdf", supp3.1, width = 8, height = 12)

supp3.2 <- do.call(plot_grid, c(lapply(rownames(logcounts(altExp(sce_split))), function(adt_name){
  ggplot(as.data.frame(logcounts(altExp(sce_split))[adt_name, altExp(sce_split)$condition_group == "CAR33-KLRC1.ko-NK_BeforeCoc"]), 
         aes_string(x=logcounts(altExp(sce_split))[adt_name, altExp(sce_split)$condition_group == "CAR33-KLRC1.ko-NK_BeforeCoc"])) +
    geom_histogram(color="black",
                   fill = "black",
                   breaks=seq(0, 4.5, by=0.10)) + 
    xlab(adt_name) + 
    theme(axis.title.x = element_text(size = 8, face="bold"))}), ncol = 4))

ggsave("fig3supp3.2.tiff", supp3.2, width = 8, height = 12)
ggsave("fig3supp3.2.pdf", supp3.2, width = 8, height = 12)

supp3.3 <- do.call(plot_grid, c(lapply(rownames(logcounts(altExp(sce_split))), function(adt_name){
  ggplot(as.data.frame(logcounts(altExp(sce_split))[adt_name, altExp(sce_split)$condition_group == "CAR33-NK_AfterCoc"]), 
         aes_string(x=logcounts(altExp(sce_split))[adt_name, altExp(sce_split)$condition_group == "CAR33-NK_AfterCoc"])) +
    geom_histogram(color="black",
                   fill = "black",
                   breaks=seq(0, 4.5, by=0.10)) + 
    xlab(adt_name) + 
    theme(axis.title.x = element_text(size = 8, face="bold"))}), ncol = 4))

ggsave("fig3supp3.3.tiff", supp3.3, width = 8, height = 12)
ggsave("fig3supp3.3.pdf", supp3.3, width = 8, height = 12)

supp3.4 <- do.call(plot_grid, c(lapply(rownames(logcounts(altExp(sce_split))), function(adt_name){
  ggplot(as.data.frame(logcounts(altExp(sce_split))[adt_name, altExp(sce_split)$condition_group == "CAR33-NK_BeforeCoc"]), 
         aes_string(x=logcounts(altExp(sce_split))[adt_name, altExp(sce_split)$condition_group == "CAR33-NK_BeforeCoc"])) +
    geom_histogram(color="black",
                   fill = "black",
                   breaks=seq(0, 4.5, by=0.10)) + 
    xlab(adt_name) + 
    theme(axis.title.x = element_text(size = 8, face="bold"))}), ncol = 4))

ggsave("fig3supp3.4.tiff", supp3.4, width = 8, height = 12)
ggsave("fig3supp3.4.pdf", supp3.4, width = 8, height = 12)
```

# Session

```{r sessionInfo, cache = 0}
date()
sessionInfo()

knitr::knit_exit()
```

